---
title: "Team Notebook"
author: "Aidan Hughes, Aya Hussein, Caley Fox Shannon"
date: "`r Sys.Date()`"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

For our data analysis project, we have chosen to look at House of Representatives Office Expenditures. We think that the dataset is rich with newsworthy details and we know that the public is generally interested in how lawmakers spend their tax dollars.


## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}

# Turn off scientific notation. 

options(scipen=999)

# Load libraries. 

library(tidyverse)
library(readr)
library(scales)
library(janitor)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}

# Load the data from ProPublica.

q4_2022 <- read_csv("data/2022Q4.csv")
q3_2022 <- read_csv("data/2022Q3.csv")
q2_2022 <- read_csv("data/2022Q2.csv")
q1_2022 <- read_csv("data/2022Q1.csv")

q4_2021 <- read_csv("data/2021Q4.csv")
q3_2021 <- read_csv("data/2021Q3.csv")
q2_2021 <- read_csv("data/2021Q2.csv")
q1_2021 <- read_csv("data/2021Q1.csv")

q4_2020 <- read_csv("data/2020Q4.csv")
q3_2020 <- read_csv("data/2020Q3.csv")
q2_2020 <- read_csv("data/2020Q2.csv")
q1_2020 <- read_csv("data/2020Q1.csv")

q4_2019 <- read_csv("data/2019Q4.csv")
q3_2019 <- read_csv("data/2019Q3.csv")
q2_2019 <- read_csv("data/2019Q2.csv")
q1_2019 <- read_csv("data/2019Q1.csv")

q4_2018 <- read_csv("data/2018Q4.csv")
q3_2018 <- read_csv("data/2018Q3.csv")
q2_2018 <- read_csv("data/2018Q2.csv")
q1_2018 <- read_csv("data/2018Q1.csv")

# Clean required data and prepare for analysis if needed. 



# Bind dataframes into one combined dataframe  

house_spend_18_23 <- bind_rows(list(q4_2022, q3_2022, q2_2022, q1_2022, q4_2021, q3_2021, q2_2021, q1_2021, q4_2020, q3_2020, q2_2020, q1_2020, q4_2019, q3_2019, q2_2019, q1_2019, q4_2018, q3_2018, q2_2018, q1_2018))





```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

